### 2019-01-31 texs/paper_base.tex
syoko
fix mathrepr

\documentclass[senior,final,11pt]{iscs-thesis}
%論文の種類とフォントサイズをオプションに
%\usepackage{graphicx}% 必要に応じて
%\usepackage{mysettings}% 自分用設定
%-------------------
\etitle{Neural Network based Decompiler for automatically recovering source codes from compiled binary data}
\jtitle{ニューラルネットを用いた逆コンパイラに}

@@

\newcommand{\argmax}{\mathop{\rm arg\,max}\limits}
\newcommand{\argmin}{\mathop{\rm arg\,smin}\limits}

@@

In this chapter, we explain about the decompiler, its statistical formalization and 
the neural network models used in the experiment.
\section{Decompiler}

@@

A compiler is a program which compiles a source code written in language Y to another language X. 
% 文献 Usually, X is a high-level language and Y is relatively a low-level language. 
For example, a GNU C compiler translates the C language to the x64 assembly language, a javac compiler translates Java to a Java virtual machine code.
% This process is usually 
% 文献
% Especially, 
% This reverse process is usually insufficient or impossible.
A decompiler is a program which aims to reverse the process of a compiler, retrieving a source code of the high-level language Y from a source code of a low-level language X. 
When Y is a high-level language and X is relatively a low-level language, 
this reverse process becomes insufficient or impossible
\citep{hex_rays,decompile_hard_java}.
% for instance X is the C language and Y is the x64 assembly language 
This is because many information, such as variable names or function names, are lost when compiling.
Additionally, the program structure described in the high level-language is lost too. 
For example, a simple \texttt{for} statement can be represented by the combination of a \texttt{while} statement and an \texttt{if} statement, or the combination of \texttt{if} statement and \texttt{goto} statement. 
So a decompiler cannot distinguish the difference in their representations, 
although in the real situation \texttt{for} statement are often used and \texttt{goto} statement are rarely used.

@@

\begin{figure}
	\begin{tabular}{cc}
		\begin{minipage}[c]{0.5\hsize}
		\fbox{
			\lstinputlisting[breaklines=false]{insert_list.c}
			}
			\caption*{Sample C source code}
		\end{minipage}
		\\ \\
		\begin{minipage}[b]{0.5\hsize} %title={Result of the Hopper Decompiler}
			\begin{framed}
			\lstinputlisting[]{snowman.c}
			\end{framed}
			\caption*{Result of the Snowman Decompiler}
		\end{minipage}
		\begin{minipage}[b]{0.5\hsize}
			\begin{framed}
			\lstinputlisting[breaklines=false]{recstudio.c}
			\end{framed}
			\caption*{Result of the REC Studio Decompiler}
		\end{minipage}
	\end{tabular}
	% \vspace*{-0.6cm}
	\caption{Decompilation results of existing decompilers for C source code}
	\label{fig:cw}
\end{figure}
Figure~\ref{fig:cw} shows an example of decompilation results of the Snowman decompiler version v0.1.3 and the REC Studio decompiler version 4.1 
for a sample C language code. The sample C code is compiled by the GNU C Compiler version 4.8.5, with the optimization level \texttt{-O0}.
The REC decompiler missed the structure of the variable \texttt{root}.
It also failed to estimate the number of arguments and the return value of the function.
The Snowman Decompiler succeeded to reconstruct the structure of the variable as \texttt{struct s0}.
It also succeeded to estimate the number of arguments and the return value as the semantically correct expressions, 
however the statement \texttt{return 0;} is reconstructed to the complex combination of statements.

These incorrect or verbose results might be due to missing patterns in the assembly codes. 
Existing decompilers are made up with many steps similarly to compilers \citep{decompiler_path,hex_rays}.
In each step, a decompiler finds patterns in binary data, converts them to more high-level structure.
However, covering all patterns might be hardly impossible, since there are many compilers and compiler optimizations \citep{hex_rays}.

By utilizing source codes corpus, a statistical method might find patterns automatically, without enumerating each patterns.
Therefore, we tried to apply some statistical methods, which are described in the next section.

@@

% 189
In this thesis, we try to apply statistical approach, specifically SMT (statistical machine translation) technique, for decompilation.
SMT is the task which tries to translate one sentence $sy$ written in one language into the sentence $sx$ written in another language.

The decompilation problem is formulated as follows. 
Suppose that there is a compiler which converts a source code $ sy $ written in a programming language 
to a source code $ sx $  written in another programming language with probability $p(sx|sy)$.
In addition, suppose that $ sy $ is human-generated source code and follows the probability distribution $ p(sy) $.
In this situation, the decompilation probability distribution $ p(sy|sx) $ is in proportion to $ p(sx|sy)p(sy) $ according to the Bayesian low, 
and the $\argmax_{sy} p(sy|sx)$ is considered as the best human-intelligible decompilation result for $sx$.

In the experiment, we collected source codes from open source projects as the random sampling from the distribution $ p(sy) $, 
and compiled them to generate the data pairs $ {\{(sx_k,sy_k)\}}_k $.
Then the data pairs are used for training the SMT models to estimate $p(sy|sx)$, and testing the performance of the models.

% Suppose that there is a compiler  Given $ sx $ as a source code of domain language X, compiler generates $sy$ as a code of target language Y with probability $p(sy|sx)$.

% Given $sx = [sx_1, \dots, sx_n] $ as a tokenized sequence of the source code of domain language X, compiler generates $sy$ as a code of target language Y with probability $p(sy|sx)$.
% Then, the best decompilation result for target source code $sy$ is $ \argmax_{sx} p(sx|sy)$. 
% To generate $ \argmax_{sx} p(sx|sy)$, we decomposit $ p(sx|sy) = p([sx_1, \dots, sx_n] |sy) = \prod_{k=1}^{n} p(sx_k|sy,[sx_1,\dots,sx_{k-1}]) $,
% where $ p(sx_k|sy,[sx_1,\dots,sx_{k-1}]) $ represents the probability that the token $ sx_k $ follows to the sequence $ [sx_1,\dots,sx_{k-1}] $ 
% when the input sequence is $ sy $. 
% We try to estimate the probability $ p(sx_k|sy,[sx_1,\dots,sx_{k-1}]) $ by LSTM, which is described in this section.
% With estimated $ p(sx_k|sy,[sx_1,\dots,sx_{k-1}]) $, we can generate approximated $ \argmax_{sx} p(sx|sy)$ sequentially from $sx_1$ to $sx_n$ by selecting token $ \argmax_{sx_{k}} p(sx_k|sy,[sx_1,\dots,sx_{k-1}]) $ at each timing.

% The LSTM is 
% We assume the prior distribution $p(sx)$ as the human-generated source code. 
% With that assumption, is in proportion to $p(sy|sx)p(sx)$ according to the Bayesian low, and 
% the $\argmax_{sx} p(sx|sy)$ is considered as the best human-intelligible decompilation result for the low-level code $sy$.  
% So we try to estimate $ \argmax_{sx} p(sy|sx)p(sx)$ for decompilation result.
% And if the compiler is deterministic, there would be a function $f$ which represents the compiler and $p(f(sx)|sx) = 1$,
% then the objective becomes to $ \argmax_{sx,f(sx)=sy} p(sx)$.
% The distribution $p(sx)$ is approximated by collecting source code from open source projects.
% So if we had enough high-level source code data, we could choose the most popular high-level source code which generates given low-level code for the decompile result.
% However, high-level source code data can't be comprehensive.
% So we have to model the structure of compilation process somehow.
% We modeled the structure by the LSTM, which is commonly used in resent SMT techniques.

% (TODO)LSTM,Seq2seqの説明、その後Attentionについて説明、さらにその後seq2treeについて説明する。
\subsection{RNN and LSTM}
The RNN (Recurrent Neural Network) is a neural network for dealing with sequence data.
The structure of the RNN is the sequence of the same node network units, as shown in figre~\ref{fig:encoder} and figre~\ref{fig:decoder}. 
The LSTM (Long Short-Term Memory) is a unit of the RNN which we used in the experiment. 
It was first introduced by \citet{first_LSTM}, for overcoming the vanishing gradient problem and exploding gradient problem.
In short, a LSTM unit receives two input vectors, 
the one is the internal state of the LSTM unit, and the other is the input for the LSTM unit, 
and the LSTM unit emits one output vectors, which represents the next internal state of the LSTM unit.

@@

The sequence-to-sequence model is the model for SMT task, introduced by \citet{seq2seq}.
% The combination of RNN and LSTM is applied to SMT task in, as the essential unit of the 
Sequence-to-sequence model is composed of two recurrent neural networks, the encoder network and the decoder network.

@@

Figure~\ref{fig:encoder} shows the structure of the encoder network.  The purpose of the encoder network is to summarize the input sequence $sx$, whose length varies depending on the sequence $sx$, into one fixed size vector.
This process is carried out as follows. 
First, $sx$ is tokenized to sequence $sx_{1} \dots sx_{n}$, and each token $sx_{k}$ is embedded to its vector representation $vx_{k}$. 
Then, the initial LSTM internal state $enc_0$ is initialized randomly, 
and $ enc_{1} \dots enc_{n} $ are calculated in accordance with the recursive formula $ enc_{k} = lstm_{enc}(enc_{k-1},vx_{k}) $.
Finally, the last internal state $ enc_{n}$ is returned as the reduced representation of the input sequence $sx$.

Figure~\ref{fig:decoder} shows the structure of the decoder network.  The purpose of the decoder network is to estimate the distribution $ p(sy_k|sx,[sy_1,\dots,sy_{k-1}]) $ for each k. 
The training process is carried out as follows. 
First, $sy$ is tokenized, embedded to a vector representation $vy_{1} \dots vy_{m}$ in the same way as embedding $sx$.
% prepended the token $ EOS $ for indicating the end of the sequence, then 
Then, the first internal state $ dec_{1}$ is calculated from the start vector $vy_{begin}$ and encoded input $ enc_{n} = dec_{0} $ in accordance with the formula  $ dec_1 = lstm_{dec}(dec_{0},y_{begin}) $.
Then, $ dec_{2} \dots dec_{m+1} $ are calculated in accordance with the recursive formula $ dec_{k+1} = lstm_{dec}(dec_{k},vy_{k}) $.
Each decoder internal state $ dec_{k} $ is passed to the linear layer $ W_{dec} $ and then softmaxed to generate a probability vector $ py_{k} = softmax(W_{dec}(dec_{k}))$.
The probability vector $ py_{k} $ is intended to represent the probability distribution $ p(sy_k|sx,[sy_1,\dots,sy_{k-1}]) $,
and last probability vector $ py_{m+1} $ is intended to represent the probability of the end of the sequence $ p(EOS|sx,[sy_1,\dots,sy_{m}]) $.
Thus the network is trained to reduce the difference between predicted distribution $ py_{k} $ and actual distribution $ p(sy_k|sx,[sy_1,\dots,sy_{k-1}]) $.
In the experiment, we choiced cross-entropy of $ py_{k} $ and $ p(sy_k|sx,[sy_1,\dots,sy_{k-1}]) $ for loss function.

The inference from sequence $sx$ is carried out by estimating
$ \argmax_{sy} p(sy|sx) = \argmax_{sy} \prod_{k=1}^n p(sy_k|sx,[sy_1,\dots,sy_{k-1}]) $.
Calculating argmax directly is exponentially time consuming, therefore we have to approximate the argmax by calculating $sy = [sy_1, \dots sy_m]$ sequentially.
First, $sy_1$ is calculated as $ \argmax_{sy_1} p(sy_1|sx,[]) $, where distribution $ p(sy_1|sx,[]) $ is approximated by 
$ softmax(W_{dec}(dec_{1}))$. 
Then, the next token $sy_2 = \argmax_{sy_1} p(sy_2|sx,[sy_2]) $ is calculated from $ dec_{1} $ and the vector embedding of $sy_1$, and so on.
When the token $sy_{m+1}$ is the $ EOS $ token, generating process is stopped and the sequence $ [sy_1, \dots sy_m]$ is returned as the inference result.

@@

First is the Bidirectional Recurrent Neural Networks (BRNN), which is the architecture introduced by \citet{BiRNN}.
In our experiment, this architecture is used for the encoder network as follows.
In addition to the internal state $ enc_{k} $, the reverse internal state $ enc_{k}^{\gets} $ is also calculated 
in accordance with the recursive formula $ enc_{k}^{\gets} = lstm_{enc}^{\gets}(enc_{k+1}^{\gets},vx_{k}) $.
Then both of the internal states $ enc_{n} $ and $ enc_{0}^{\gets} $ are used for decoder network.

Second is the multi-layer recurrent neural networks, introduced by \citet{multi_layer}.
They reported that, feeding the output sequence of the LSTM network to another LSTM network, like stacking the network layers, will improve the accuracy of the prediction.
This architecture is called as the stacked recurrent neural networks or the multi-layer recurrent neural networks.

@@

\subsection{Attention}
Attention is a extension module for sequence-to-sequence model to improve the quality of translation. 
A simple sequence-to-sequence model has a problem that the size of the encoded vector $enc_{n}$ is fixed 
regardless of the length of the input sequence $sx$. 
Therefore, the longer the length of the input becomes, the more likely the information of the input will be lost.
This problem is eased by using the information in the encoding network when decoding the output sequence.

The attention module is first developed by \citet{attention_paper}.
We implemented the global general attention, which is introduced by \citet{dot_attention}, for improving the quality of decompilation.
In the following, we will describe about the global general attention.

As the mentioned in the previous section, the pure sequence-to-sequence model uses the output of the decoder network $ dec_{k} $ for estimating 
the distribution $ p(sy_k|sx,[sy_1,\dots,sy_{k-1}]) $.
In the sequence-to-sequence model with attention, 
a context vector $ ctx_{k} $ is calculalted from $ dec_{k} $ and the internal states of encoder network $ enc_{1} \dots enc_{n} $, 
then the distribution is estimated from the output vector $ dec_{k} $ and context vector $ ctx_{k} $.
The context vector $ ctx_{k} $ is calculated as follows.
For each encoder state $ enc_{l} $, the importance of the state is calculated as $ a_{l} = score(dec_{k},enc_{l}) $,
then the $ ctx_{k} $ is calculated as $ ctx_{k} = \sum_{l=1}^{n} \frac{\exp{a_{l}}}{\sum_{l=1}^{n}\exp{a_{l}}} enc_{l} $, 
as the weighted average of $ enc_{1} \dots enc_{n} $.
% Multiple score functions are introduced by \cite{dot_attention}, and we choiced 
In the paper by \citet{dot_attention}, various kinds of score function was suggested.
In our experiment, we used general score function , which is calculated as $ score(dec_{k},enc_{l}) = dec_{k}^T W enc_{l} $ ,
where W is the parameter matrix.

% With the information of the internal state $ ctx_{k} $ of the encoder network, the decoder network can access to the

% アテンション機構を用いる場合、前述したようにdecoder networkの出力 $ dec_{k} $ をそのまま用いて分布をesitmateするのではなく、
% $ dec_{k} $ と encoder network の出力 $ enc_{1} \dots enc_{n} $ (encoder が BiLSTMの場合 $ enc^{\gets}_{1} \dots enc^{\gets}_{n} $ も)
% を用いて context vector $ ctx_{k} $ を生成し、 $ dec_{k} $ と $ ctx_{k} $ によって条件付き分布をesitmateする。

% $ ctx_{k} $ は、 $ enc_{1} \dots enc_{n} $ の加重平均によって計算される。 具体的には、
% 各状態 $ enc_{l} $ について その状態のスコア $ a_{l} = score(dec_{k},enc_{l}) $ を計算し、
% $ ctx_{k} = \sum_{l=1}^{n} \frac{\exp{a_{l}}}{\sum_{l=1}^{n}\exp{a_{l}}} enc{l} $ として計算される。

% この $ ctx_{k} $ を併せて用いることにより、 分布予測時に入力データの情報にアクセスすることができ、精度の向上が見込める。

% In the pure sequence-to-sequence model, the conditional probability  $ p(sy_k|sx,[sy_1,\dots,sy_{k-1}]) $ is calculated from $ dec_{k} $ ,which is the internal state of the decoder network.
% In the sequence-to-sequence with attention model, the probability is calcutated from $ dec_{k} $ and the context vevctor $ c_{k} $.
% $ c_{k} $

@@

% The attentional hidden state $ dec'_{k} $ is calculated from the $ dec_{k} $ and $ c_{k} $.
% Then, the probability $ p(sy_k|sx,[sy_1,\dots,sy_{k-1}]) $ is estimated as $ softmax(W_{dec}(dec'_{k}))$, instead of $ softmax(W_{dec}(dec_{k}))$.

\subsection{Sequence-to-Tree}
% 184 words
Compare to natural languages, programming languages obeys rigid grammars. % wich are usually context free grammers.
Therefore, it is reasonable to consider the grammar of the programming language when generating programs.
Sequence-to-Tree is a model for generating a tree-structured output from a sequence input. 
This model is introduced by \citet{Seq2Tree} for generating a program from a program specification written in a natural language. 
This task is similar to our task which aims to generate programs from the assembly codes, which are a sequence of instructions.
Therefore we implemented this model in the experiment.

Now, we will describe about the sequence-to-tree model.
sequence-to-tree model is composed of two networks, the encoder network and the decoder network.
The encoder network of the sequence-to-tree model is the same as that of the sequence-to-seequence model. It summarizes the input sequence into one fixed size vector.

\begin{figure}[]
	\begin{center}
	\includegraphics[height=11.5cm]{ast_zu.png}
	\end{center}
	\caption{AST representation of a C statement \texttt{b = a + 3;} }
	\label{fig:ast_zu}
\end{figure}

>>>>

\documentclass[senior,final,11pt]{iscs-thesis}
%論文の種類とフォントサイズをオプションに
%\usepackage{graphicx}% 必要に応じて
%\usepackage{mysettings}% 自分用設定
%-------------------
\etitle{Neural Network based Decompiler and its Evaluation}
\jtitle{ニューラルネットを用いた逆コンパイラとその評価}

@@

\newcommand{\argmax}{\mathop{\rm arg\,max}\limits}
%\newcommand{\argmin}{\mathop{\rm arg\,smin}\limits}

@@

In this chapter, we explain a decompiler, its statistical formalization and 
the neural network models used in the experiment.
\section{Decompiler}

@@

A compiler is a program which compiles a source code written in language Y to another language X. 
% 文献 Usually, X is a high-level language and Y is relatively a low-level language. 
For example, a GNU C compiler translates the C language to the x64 assembly language, a javac compiler translates Java to a Java virtual machine code.
% This process is usually 
% 文献
% Especially, 
% This reverse process is usually insufficient or impossible.
A decompiler is a program which aims to reverse the process of a compiler, retrieving a source code of the high-level language Y from a source code of a low-level language X. 
When Y is a high-level language and X is relatively a low-level language, 
this reverse process becomes insufficient or impossible
\citep{hex_rays,decompile_hard_java}.
% for instance X is the C language and Y is the x64 assembly language 
This is because a lot of information, such as variable names or function names, is lost when compiling.
Additionally, the program structure described in the high level-language is lost too. 
For example, a simple \texttt{for} statement can be represented by the combination of a \texttt{while} statement and an \texttt{if} statement, or the combination of an \texttt{if} statement and a \texttt{goto} statement. 
So a decompiler cannot distinguish the difference in their representations, 
although in the real situation \texttt{for} statements are often used and \texttt{goto} statements are rarely used.

@@

\begin{figure}
	\begin{tabular}{cc}
		\begin{minipage}[c]{0.5\hsize}
		\fbox{
			\lstinputlisting[breaklines=false]{insert_list.c}
			}
			\caption*{Sample C source code}
		\end{minipage}
		\\ \\
		\begin{minipage}[b]{0.5\hsize} %title={Result of the Hopper Decompiler}
			\begin{framed}
			\lstinputlisting[]{snowman.c}
			\end{framed}
			\caption*{Result of the Snowman Decompiler}
		\end{minipage}
		\begin{minipage}[b]{0.5\hsize}
			\begin{framed}
			\lstinputlisting[breaklines=false]{recstudio.c}
			\end{framed}
			\caption*{Result of the REC Studio Decompiler}
		\end{minipage}
	\end{tabular}
	% \vspace*{-0.6cm}
	\caption{Decompilation results of existing decompilers for C source code}
	\label{fig:cw}
\end{figure}
Figure~\ref{fig:cw} shows an example of decompilation results of the Snowman decompiler version v0.1.3 and the REC Studio decompiler version 4.1 
for a sample C language code. The sample C code is compiled by the GNU C Compiler version 4.8.5, with the optimization level \texttt{-O0}.
The REC decompiler missed the structure of the variable \texttt{root}.
It also failed to estimate the number of arguments and the return value of the function.
The Snowman Decompiler succeeded in reconstructing the structure of the variable as \texttt{struct s0}.
It also succeeded in estimating the number of arguments and the return value as semantically correct expressions. 
However the statement \texttt{return 0;} is reconstructed to a complex combination of statements.

These incorrect or verbose results might be due to missing patterns in the assembly codes. 
Existing decompilers are made up with many steps similarly to compilers \citep{decompiler_path,hex_rays}.
In each step, a decompiler finds patterns in binary data, converts them to more high-level structure.
However, covering all patterns might be hardly possible, since there are many compilers and compiler optimizations \citep{hex_rays}.

By utilizing source code corpora, a statistical method might find patterns automatically, without enumerating each pattern.
Therefore, we will apply some statistical methods, which are described in the next section.

@@

% 189
In this thesis, we try to apply a statistical approach, specifically a statistical machine translation (SMT) technique, for decompilation.
SMT is the task which tries to translate one sentence $y$ written in one language into the sentence $x$ written in another language.

The decompilation problem is formulated as follows. 
Suppose that there is a compiler which converts a source code $ y $ written in one programming language 
to a source code $ x $  written in another programming language with probability $p(x|y)$.
In addition, suppose that $ y $ is a human-generated source code and follows the probability distribution $ p(y) $.
In this situation, the decompilation probability distribution $ p(y|x) $ is in proportion to $ p(x|y)p(y) $ according to the Bayes law, 
and $\argmax_{y} p(y|x)$ is considered as the best human-intelligible decompilation result for $x$.

In the experiment, we collected source codes from open source projects, which are regarded as the random sampling from the distribution $ p(y) $, 
and compiled them to generate the data pairs $ {\{(x_k,y_k)\}}_k $.
Then the data pairs are used for training the SMT models to estimate $p(y|x)$, and testing the performance of the models.

% Suppose that there is a compiler  Given $ x $ as a source code of domain language X, compiler generates $y$ as a code of target language Y with probability $p(sy|x)$.

% Given $x = [x_1, \dots, x_n] $ as a tokenized sequence of the source code of domain language X, compiler generates $sy$ as a code of target language Y with probability $p(sy|x)$.
% Then, the best decompilation result for target source code $sy$ is $ \argmax_{x} p(x|sy)$. 
% To generate $ \argmax_{x} p(x|sy)$, we decomposit $ p(x|sy) = p([x_1, \dots, x_n] |sy) = \prod_{k=1}^{n} p(x_k|sy,[x_1,\dots,x_{k-1}]) $,
% where $ p(x_k|sy,[x_1,\dots,x_{k-1}]) $ represents the probability that the token $ x_k $ follows to the sequence $ [x_1,\dots,x_{k-1}] $ 
% when the input sequence is $ sy $. 
% We try to estimate the probability $ p(x_k|sy,[x_1,\dots,x_{k-1}]) $ by LSTM, which is described in this section.
% With estimated $ p(x_k|sy,[x_1,\dots,x_{k-1}]) $, we can generate approximated $ \argmax_{x} p(x|sy)$ sequentially from $x_1$ to $x_n$ by selecting token $ \argmax_{x_{k}} p(x_k|sy,[x_1,\dots,x_{k-1}]) $ at each timing.

% The LSTM is 
% We assume the prior distribution $p(x)$ as the human-generated source code. 
% With that assumption, is in proportion to $p(sy|x)p(x)$ according to the Bayesian low, and 
% the $\argmax_{x} p(sx|y)$ is considered as the best human-intelligible decompilation result for the low-level code $y$.  
% So we try to estimate $ \argmax_{sx} p(y|sx)p(sx)$ for decompilation result.
% And if the compiler is deterministic, there would be a function $f$ which represents the compiler and $p(f(sx)|sx) = 1$,
% then the objective becomes to $ \argmax_{sx,f(sx)=y} p(sx)$.
% The distribution $p(sx)$ is approximated by collecting source code from open source projects.
% So if we had enough high-level source code data, we could choose the most popular high-level source code which generates given low-level code for the decompile result.
% However, high-level source code data can't be comprehensive.
% So we have to model the structure of compilation process somehow.
% We modeled the structure by the LSTM, which is commonly used in resent SMT techniques.

% (TODO)LSTM,Seq2seqの説明、その後Attentionについて説明、さらにその後seq2treeについて説明する。
\subsection{RNN and LSTM}
A recurrent neural network (RNN) is a neural network for dealing with sequence data.
The structure of the RNN is the sequence of the same node network units, as shown in Figre~\ref{fig:encoder} and Figre~\ref{fig:decoder}. 
The long short-term memory (LSTM) is a unit of the RNN which we used in the experiment. 
It was first introduced by \sloppy \citet{first_LSTM} , for overcoming the vanishing gradient problem and the exploding gradient problem.
In short, an LSTM unit receives two input vectors, 
one is the internal state of the LSTM unit, and the other is the input for the LSTM unit, 
and the LSTM unit emits one output vector, which represents the next internal state of the LSTM unit.

@@

The sequence-to-sequence model is the model for SMT task, introduced by \citet{seq2seq}.
% The combination of RNN and LSTM is applied to SMT task in, as the essential unit of the 
The sequence-to-sequence model is composed of two recurrent neural networks, the encoder network and the decoder network.

@@

Figure~\ref{fig:encoder} shows the structure of the encoder network.  The purpose of the encoder network is to summarize the input sequence $x$, whose length varies depending on the sequence $x$, into a vector of fixed size.
This process is carried out as follows. 
First, $x$ is tokenized to sequence $x_{1} \dots x_{n}$, and each token $x_{k}$ is embedded to its vector representation $\mathrm{vx}_{k}$. 
Then, the initial LSTM internal state $\mathrm{enc}_0$ is initialized randomly, 
and $ \mathrm{enc}_{1} \dots \mathrm{enc}_{n} $ are calculated in accordance with the recursive formula $ \mathrm{enc}_{k} = \mathrm{lstm}_{\mathrm{enc}}(\mathrm{enc}_{k-1},\mathrm{vx}_{k}) $.
Finally, the last internal state $ \mathrm{enc}_{n}$ is returned as the reduced representation of the input sequence $x$.

Figure~\ref{fig:decoder} shows the structure of the decoder network.  The purpose of the decoder network is to estimate the distribution $ p(y_k|x,[y_1,\dots,y_{k-1}]) $ for each $ k $. 
The training process is carried out as follows. 
First, $y$ is tokenized, embedded to a vector representation $\mathrm{vy}_{1} \dots \mathrm{vy}_{m}$ in the same way as embedding $x$.
% prepended the token $ EOS $ for indicating the end of the sequence, then 
Then, the first internal state $ \mathrm{dec}_{1}$ is calculated from the start vector $\mathrm{vy}_{\mathrm{begin}}$ and encoded input $ \mathrm{enc}_{n} = \mathrm{dec}_{0} $ in accordance with the formula  $ \mathrm{dec}_1 = \mathrm{lstm}_{\mathrm{dec}}(\mathrm{dec}_{0},\mathrm{vy}_{\mathrm{begin}}) $.
Then, $ \mathrm{dec}_{2} \dots \mathrm{dec}_{m+1} $ are calculated in accordance with the recursive formula $ \mathrm{dec}_{k+1} = \mathrm{lstm}_{\mathrm{dec}}(\mathrm{dec}_{k},\mathrm{vy}_{k}) $.
Each decoder internal state $ \mathrm{dec}_{k} $ is passed to the linear layer $ W_{\mathrm{dec}} $ and then softmaxed to generate a probability vector $ \mathrm{py}_{k} = \mathrm{softmax}(W_{\mathrm{dec}}(\mathrm{dec}_{k}))$.
The probability vector $ \mathrm{py}_{k} $ is intended to represent the probability distribution $ p(y_k|x,[y_1,\dots,y_{k-1}]) $,
and last probability vector $ \mathrm{py}_{m+1} $ is intended to represent the probability of the end of the sequence $ p(\mathrm{EOS}|x,[y_1,\dots,y_{m}]) $.
Thus the network is trained to reduce the difference between predicted distribution $ \mathrm{py}_{k} $ and actual distribution $ p(y_k|x,[y_1,\dots,y_{k-1}]) $.
In the experiment, we chose cross-entropy of $ \mathrm{py}_{k} $ and $ p(y_k|x,[y_1,\dots,y_{k-1}]) $ as a loss function.

The inference from sequence $x$ is carried out by estimating
$ \argmax_{y} p(y|x) = \argmax_{y} \prod_{k=1}^n p(y_k|x,[y_1,\dots,y_{k-1}]) $.
Calculating argmax directly is exponentially time consuming, therefore we have to approximate the argmax by calculating $y = [y_1, \dots, y_m]$ sequentially.
First, $y_1$ is calculated as $ \argmax_{y_1} p(y_1|x,[]) $, where the distribution $ p(y_1|x,[]) $ is approximated by 
$ \mathrm{softmax}(W_{\mathrm{dec}}(\mathrm{dec}_{1}))$. 
Then, the next token $y_2 = \argmax_{y_2} p(y_2|x,[y_1]) $ is calculated from $ \mathrm{dec}_{1} $ and the vector embedding of $y_1$, and so on.
When the token $y_{m+1}$ is the EOS token, the generating process is stopped and the sequence $ [y_1, \dots ,y_m]$ is returned as an inference result.

@@

The first is the bidirectional RNN (BRNN), which is the architecture introduced by \citet{BiRNN}.
In our experiment, this architecture is used for the encoder network as follows.
In addition to the internal state $ \mathrm{enc}_{k} $, the reverse internal state $ \mathrm{enc}_{k}^{\gets} $ is also calculated 
in accordance with the recursive formula $ \mathrm{enc}_{k}^{\gets} = \mathrm{lstm}_{\mathrm{enc}}^{\gets}(\mathrm{enc}_{k+1}^{\gets},\mathrm{vx}_{k}) $.
Then both of the internal states $ \mathrm{enc}_{n} $ and $ \mathrm{enc}_{0}^{\gets} $ are used for the decoder network.

The second is the multi-layer RNN, introduced by \citet{multi_layer}.
They reported that, feeding the output sequence of one LSTM network to another LSTM network, similarly to stacking the network layers, will improve the accuracy of prediction.
This architecture is called the stacked RNN or the multi-layer RNN.

@@

\subsection{Attention}
Attention is an extension module for the sequence-to-sequence model to improve the quality of translation. 
A simple sequence-to-sequence model has a problem that the size of the encoded vector $enc_{n}$ is fixed 
regardless of the length of the input sequence $x$. 
Therefore, the longer the length of the input becomes, the more likely the information of the input will be lost.
This problem is eased by using the information in the encoding network when decoding the output sequence.

The attention module was first developed by \citet{attention_paper}.
We implemented the global general attention, which was introduced by \citet{dot_attention}, for improving the quality of decompilation.
In the following, we will describe the global general attention.

As the mentioned in the previous section, the pure sequence-to-sequence model uses the output of the decoder network $ \mathrm{dec}_{k} $ for estimating 
the distribution $ p(y_k|x,[y_1,\dots,y_{k-1}]) $.
In the sequence-to-sequence model with attention, 
a context vector $ \mathrm{ctx}_{k} $ is calculalted from $ \mathrm{dec}_{k} $ and the internal states of encoder networks $ \mathrm{enc}_{1} \dots \mathrm{enc}_{n} $, 
then the distribution is estimated from the output vector $ \mathrm{dec}_{k} $ and context vector $ \mathrm{ctx}_{k} $.
The context vector $ \mathrm{ctx}_{k} $ is calculated as follows.
For each encoder state $ \mathrm{enc}_{l} $, the importance of the state is calculated as $ a_{l} = \mathrm{score}(\mathrm{dec}_{k},\mathrm{enc}_{l}) $,
then $ \mathrm{ctx}_{k} $ is calculated as $ \mathrm{ctx}_{k} = \sum_{l=1}^{n} \frac{\exp{a_{l}}}{\sum_{l=1}^{n}\exp{a_{l}}} \mathrm{enc}_{l} $, 
as the weighted average of $ \mathrm{enc}_{1} \dots \mathrm{enc}_{n} $.
% Multiple score functions are introduced by \citet{dot_attention}, and we choiced 
In \citet{dot_attention}, various kinds of score functions were suggested.
In our experiment, we used a general score function, which is calculated as $ \mathrm{score}(\mathrm{dec}_{k},\mathrm{enc}_{l}) = \mathrm{dec}_{k}^T W \mathrm{enc}_{l} $ ,
where $ W $ is a parameter matrix.

% With the information of the internal state $ \mathrm{ctx}_{k} $ of the encoder network, the decoder network can access to the

% アテンション機構を用いる場合、前述したようにdecoder networkの出力 $ dec_{k} $ をそのまま用いて分布をesitmateするのではなく、
% $ dec_{k} $ と encoder network の出力 $ enc_{1} \dots enc_{n} $ (encoder が BiLSTMの場合 $ enc^{\gets}_{1} \dots enc^{\gets}_{n} $ も)
% を用いて context vector $ \mathrm{ctx}_{k} $ を生成し、 $ dec_{k} $ と $ \mathrm{ctx}_{k} $ によって条件付き分布をesitmateする。

% $ \mathrm{ctx}_{k} $ は、 $ enc_{1} \dots enc_{n} $ の加重平均によって計算される。 具体的には、
% 各状態 $ enc_{l} $ について その状態のスコア $ a_{l} = score(dec_{k},enc_{l}) $ を計算し、
% $ \mathrm{ctx}_{k} = \sum_{l=1}^{n} \frac{\exp{a_{l}}}{\sum_{l=1}^{n}\exp{a_{l}}} enc{l} $ として計算される。

% この $ \mathrm{ctx}_{k} $ を併せて用いることにより、 分布予測時に入力データの情報にアクセスすることができ、精度の向上が見込める。

% In the pure sequence-to-sequence model, the conditional probability  $ p(y_k|x,[y_1,\dots,y_{k-1}]) $ is calculated from $ dec_{k} $ ,which is the internal state of the decoder network.
% In the sequence-to-sequence with attention model, the probability is calcutated from $ dec_{k} $ and the context vevctor $ c_{k} $.
% $ c_{k} $

@@

% The attentional hidden state $ dec'_{k} $ is calculated from the $ dec_{k} $ and $ c_{k} $.
% Then, the probability $ p(y_k|x,[y_1,\dots,y_{k-1}]) $ is estimated as $ \mathrm{softmax}(W_{dec}(dec'_{k}))$, instead of $ \mathrm{softmax}(W_{dec}(dec_{k}))$.

\subsection{Sequence-to-Tree}
% 184 words
Compared to natural languages, programming languages obey rigid grammars. % wich are usually context free grammers.
Therefore, it is reasonable to consider the grammar of the programming language when generating programs.
Sequence-to-tree is a model for generating a tree-structured output from a sequence input. 
This model was introduced by \citet{Seq2Tree} for generating a program from a program specification written in natural language. 
This task is similar to our task which aims to generate programs from assembly codes, which are sequences of instructions.
Therefore we implemented this model in the experiment.

Now, we will describe the sequence-to-tree model.
The sequence-to-tree model is composed of two networks, the encoder network and the decoder network.
The encoder network of the sequence-to-tree model is the same as that of the sequence-to-seequence model. 
It summarizes an input sequence into a vector of fixed size.

\begin{figure}[]
	\begin{center}
	\includegraphics[height=11.5cm]{ast_zu.png}
	\end{center}
	\caption{abstract syntax tree (AST) representation of a C statement \texttt{b = a + 3;} }
	\label{fig:ast_zu}
\end{figure}
